Alternatives for Neighborhood Function in Kohonen Maps \thanks{This work was supported by a private funding of Velbazhd Software LLC.}

### 2019-02-24 Alternatives-for-Neighborhood-Function-in-Kohonen-Maps/paper.tex
Small additions.

Self-organizing maps or Kohonen Neural Networks (KNNs) are networks with unsupervised training \cite{schreck01}. They are very useful in finding nonlinear dependencies when data are presented in very high dimensional spaces. The projection from the high-dimensional space is usually done in a rectangular lower-dimensional space. The main idea behind KNNs is the organization of unlabeled vectors with particular features in predefined number of groups called clusters \cite{cottrell01}. Grid of the self-organizing map is a handy tool for convenient visualization which can reveal different features of the network. SOMs are attractive when there consist of two or more separate regions. The real goal is not to find a perfect clustering but get good idea of the cluster structure.

@@

Generally accepted concept of an optimal clustering is data set separation which minimizes the distance inside the cluster and maximizes the distance between the clusters. The most used form of distance measurement is the Euclidean norm.

@@

The network training is done iteratively. A data sample vector is randomly chosen from the input data set, as first step. As second step, distances between the selected vector and all the prototype vectors are computed. The best matching unit is the grid unit with prototype closest the selected vector. In the third step, the prototype vectors in the grid are updated. The best matching unit and its topological neighbors are moved closer to the input vector in the input space \cite{vesanto01}.

@@

@@

@@

>>>>

Self-organizing maps or Kohonen Neural Networks (KNNs) are networks with unsupervised training \cite{schreck01}. They are very useful in finding nonlinear dependencies when data are presented in very high dimensional spaces. The projection from the high-dimensional space is usually done in a rectangular lower-dimensional space \cite{macq01}. The main idea behind KNNs is the organization of unlabeled vectors with particular features in predefined number of groups called clusters \cite{cottrell01}. Grid of the self-organizing map is a handy tool for convenient visualization which can reveal different features of the network. SOMs are attractive when there consist of two or more separate regions. The real goal is not to find a perfect clustering but get good idea of the cluster structure.

@@

Generally accepted concept of an optimal clustering is data set separation which minimizes the distance inside the cluster and maximizes the distance between the clusters. The most used form of distance measurement is the Euclidean norm \cite{lozano01}.

@@

The network training is done iteratively. A data sample vector is randomly chosen from the input data set, as first step. As second step, distances between the selected vector and all the prototype vectors are computed. The best matching unit is the grid unit with prototype closest the selected vector. In the third step, the prototype vectors in the grid are updated. The best matching unit and its topological neighbors are moved closer to the input vector in the input space \cite{vesanto01,аmerijckx01}.

@@

\bibitem{macq01}
Macq, D., Verleysen, M., Jespers, P., Legat, J.D.: Analog implementation of a Kohonen map with on-chip learning. IEEE Transactions on Neural Networks, \textbf{4}(3), 456--461 (1993)

@@

\bibitem{lozano01}
Lozano, S., Guerrero, F., Onieva, L., Larraneta, J.: Kohonen maps for solving a class of location-allocation problems. European Journal of Operational Research, \textbf{108}(1), 106--117 (1998)

@@

\bibitem{аmerijckx01}
Amerijckx, C., Verleysen, M., Thissen, P., Legat, J.D.: Image compression by self-organized Kohonen map. IEEE Transactions on Neural Networks, \textbf{9}(3), 503--507 (1998)
